Computing Entropies With Nested Sampling

### 2017-03-17 paper/ms.tex
Random writing in appendix
Finished algorithm
Algorithm color

\While{more iterations desired}
    \State Generate $\x_{\rm ref}$ from $p(\x)$
            \Comment{Generate a reference point}
    \State $k \leftarrow 0$
            \Comment{Initialise counter}
  {\color{green}
    \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
            \Comment{Generate initial NS particles}
    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
            \Comment{Calculate distance of each particle from the reference
                     point}
    \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
            \Comment{Find the worst particle (greatest distance)}
    \State $d_{\rm max} \leftarrow d_{i^*}$
            \Comment{Find the greatest distance}
%%    \While{$d_{\rm max} > \epsilon$}
%%        \State Blah
%%    \EndWhile

  } % \color{green}
\State More to come...
\EndWhile
\end{algorithmic}
\caption{Calculates $H_{\x} = -\int p(\x) \log p(\x) \, d\x$
         without evaluating $\log p(\x)$.}
\end{algorithm}

@@

Continuous hypothesis spaces are commonly used in practice. Suppose
$\x$ has probability density $f(\x)$. The algorithm actually estimates
\begin{align}
Q &= \int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
                        \, d\x' \, d\x.
\end{align}
If we actually want the differential
entropy
\begin{align}
h_{\x} &= -\int f(x) \log f(x) \, dx
\end{align}
we can use the fact that density equals mass divided
by volume. Assume $\epsilon$ is small, so that
\begin{align}
P(d(\x', \x) < \epsilon | \x)
    &\approx
    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
    &= f(\x) \, V
\end{align}
where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.

@@

In practice, we may
want to ask a question that is different from the central issue.
For example, suppose we want to know the value of $x$ to within a tolerance
of $\pm 1$. Any of the following statements would resolve the issue:
\begin{itemize}
\item $x \in \{1, 2, 3\}$ and anything that implies it
\item $x \in \{2, 3, 4\}$ and anything that implies it
\item $x \in \{3, 4, 5\}$ and anything that implies it
\item and so on.
\end{itemize}

>>>>

\While{more iterations desired}
    \State $k \leftarrow 0$
            \Comment{Initialise counter}
    \State Generate $\x_{\rm ref}$ from $p(\x)$
            \Comment{Generate a reference point}
\color{green}
    \State Generate $\left\{\x_i\right\}_{i=1}^N$ from $p(\x)$
            \Comment{Generate initial NS particles}
    \State Calculate $d_i \leftarrow d(\x_i; \x_{\rm ref})$ for all $i$
            \Comment{Calculate distance of each particle from the reference
                     point}
    \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
            \Comment{Find the worst particle (greatest distance)}
    \State $d_{\rm max} \leftarrow d_{i^*}$
            \Comment{Find the greatest distance}
    \While{$d_{\rm max} > \epsilon$}
        \State Replace $\x_{i^*}$ with $\x_{\rm new}$ from
               $p\left(\x | (d(\x) < d_{\rm max})\right)$
            \Comment{Replace worst particle}
        \State Calculate $d_{i^*} \leftarrow d(\x_{\rm new}; \x_{\rm ref})$
            \Comment{Calculate distance of new particle from reference point}
        \State $i^* \leftarrow \textnormal{argmin}(\lambda i \rightarrow d_i)$
                \Comment{Find the worst particle (greatest distance)}
        \State $d_{\rm max} \leftarrow d_{i^*}$
                \Comment{Find the greatest distance}
        \State $k \leftarrow k+1$
            \Comment{Increment counter $k$}
    \EndWhile
\color{black}
    \State $\widehat{\mathbf{h}} \leftarrow \widehat{\mathbf{h}} + [k/N]$
            \Comment{Append latest estimate to results}
\EndWhile
\State $\widehat{h_{\rm final}} \leftarrow \frac{1}{\rm num\_iterations} \sum \widehat{\mathbf{h}}$
            \Comment{Average results}
\end{algorithmic}
\caption{Calculates $H_{\x} = -\int p(\x) \log p(\x) \, d\x$
         without evaluating $\log p(\x)$. Technically, what it actually
         estimates is $\int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
                        \, d\x' \, d\x$, the expected value of the
         log-probability of a small region near $\x$.
        The part highlighted in green is standard Nested Sampling.}
\end{algorithm}

@@

Continuous hypothesis spaces are commonly used in practice. Suppose
$\x$ has probability density $f(\x)$. The algorithm actually estimates
\begin{align}
h &= \int f(\x) \int f(\x')
            \log \left[ P(d(\x', \x) < \epsilon | \x) \right]
                        \, d\x' \, d\x.
\end{align}
If we actually want the differential
entropy
\begin{align}
h_{\x} &= -\int f(x) \log f(x) \, dx
\end{align}
we can use the fact that density equals mass divided
by volume. Assume $\epsilon$ is small, so that
\begin{align}
P(d(\x', \x) < \epsilon | \x)
    &\approx
    f(\x) \int_{d(\x', \x) < \epsilon} \, d\x'\\
    &= f(\x) \, V
\end{align}
where $V$ is the volume of the region where $d(\x', \x) < \epsilon$.

@@

In practice, we may
want to ask a question that is different from the central issue.
For example, suppose we want to know the value of $x$ to within a tolerance
of $\pm 1$. Any of the following statements would resolve the issue:
\begin{itemize}
\item $x \in \{1, 2, 3\}$ and anything that implies it
\item $x \in \{2, 3, 4\}$ and anything that implies it
\item $x \in \{3, 4, 5\}$ and anything that implies it
\item and so on.
\end{itemize}
To quantify the degree to which this question is unanswered by
$\top$, we can write the question as a union of ideal questions:
\begin{align}
Q &= \left[\downarrow (x \in \{1, 2, 3\})\right] \cup
     \left[\downarrow (x \in \{2, 3, 4\})\right] \cup ...
\end{align}

Its quantification is therefore
\begin{align}
H(Q) &= -\sum_{i=1}^3 p_i \log p_i - \sum_{i=2}^4 p_i \log p_i
+ \sum_{i=2}^3 p_i \log p_i
\end{align}
